{\bf Demystifying Debugging}

### 2012-11-08 gpi-debug/debugging.tex
final revisions i think

@@

	\linegap
	Sorts OK into the middle and front of list:
	\begin{itemize}
		\item \texttt{insert\_sorted} $\quad \hilight{olivegreen}{10} \quad  [20,30] \quad \Rightarrow \quad [\hilight{olivegreen}{10},20,30]$
		\item \texttt{insert\_sorted} $\quad \hilight{olivegreen}{32} \quad  [16,64] \quad \Rightarrow \quad [16,\hilight{olivegreen}{32},64]$
	\end{itemize}

	\linegap
	But not so great at the end:
	\begin{itemize}
	\item \texttt{insert\_sorted} $\quad \hilight{red}{100}  \quad [97,98,99] \quad \Rightarrow \quad [97,98,\hilight{red}{100},99]$
	\end{itemize}
\end{frame}

	\linegap
	In case you're curious, here's the bug:

	\linegap
		\texttt{~~~~insert\_sorted(list,~element)~\{} \\
			\texttt{~~~~~~~~\hilight{olivegreen}{int}~i~=~\hilight{brickred}{0};} \\
		\texttt{~~~~~~~~\hilight{brown}{while}~(i~<~length(list))~\{~\hilight{darkcyan}{//~Should~be~"<="!}} \\
		\texttt{~~~~~~~~~~~~\hilight{brown}{if}~(element~<~list[i])~\{} \\
		\texttt{~~~~~~~~~~~~~~~~\hilight{brown}{break};} \\
		\texttt{~~~~~~~~~~~~\}} \\
		\texttt{~~~~~~~~\}} \\
		\texttt{~~~~~~~~insert\_at\_index(list,~element,~i);} \\
		\texttt{~~~~\}} \\
\end{frame}

@@

\subsection{Time Travel}

@@

\begin{frame}{Asserts: Stealthier, more concealable prints}
	Adding print statements everywhere is very verbose
	\begin{itemize}
		\item Pro: Get to see everything
		\item Con: Have to look at everything
	\end{itemize}
	\pause
	15-122's C0 ``contracts'' are assertions
	\begin{itemize}
		\item You don't need to use C0 to write assert-ful code
		\item Asserts can go anywhere, not just around functions/loops
	\end{itemize}
	\pause
	\linegap

	\texttt{~~~~\hilight{brown}{if}~(something\_that\_should\_never\_happen)~\{} \\
	\texttt{~~~~~~~~print(\hilight{brickred}{"oh~no,~it~happened!"});} \\
	\texttt{~~~~~~~~crash();} \\
	\texttt{~~~~\}} \\
\end{frame}
\begin{frame}{Asserts: Stealthier, more concealable prints}
	Adding print statements everywhere is very verbose
	\begin{itemize}
		\item Pro: Get to see everything
		\item Con: Have to look at everything
	\end{itemize}
	15-122's C0 ``contracts'' are assertions
	\begin{itemize}
		\item You don't need to use C0 to write assert-ful code
		\item Asserts can go anywhere, not just around functions/loops
	\end{itemize}
	\linegap

@@

\begin{frame}{Aside: 15-213 ``malloc lab''}
	Many students think: ``I just want to write \texttt{malloc()}. I can do the heap checker later.''
	\pause
	\begin{itemize}
		\item Think of it like a C0 contract that must hold forever.
		\item \textbf{Write the heap checker simultaneously!}
			\pause
			\begin{itemize}
				\item Save time/effort debugging everything else
				\item Your debugging skill will level up
				\item Ben jumps up and down for emphasis
			\end{itemize}
	\end{itemize}
\end{frame}

\subsection{Space Travel}

@@

	\texttt{~~~~input~=~\hilight{brickred}{[99]};~~~~~\hilight{darkcyan}{//~OK}} \\
	\texttt{~~~~input~=~\hilight{brickred}{[100,99]};~\hilight{darkcyan}{//~OK}} \\
	\pause
	\texttt{~~~~input~=~\hilight{brickred}{[99,100]};~\hilight{darkcyan}{//~Fails}} \\
\end{frame}

@@

@@

\begin{frame}{Rubber Duck / Cardboard Dog Debugging}
	\begin{columns}[l]
	\column{0.75\textwidth}
	Hacker folklore says: describe your problem to a duck!
	\begin{itemize}
		\item ``What could I possibly get out of talking to an inanimate object?''
	\end{itemize}
	\linegap

@@

	Helps figure out what part of the story to start searching at
	}
	\column{0.25\textwidth}
	\includegraphics[width=\textwidth]{rubberduck.jpg}
	\end{columns}
\end{frame}
\begin{frame}{Rubber Duck / Cardboard Dog Debugging}
	\begin{columns}[l]
	\column{0.75\textwidth}
	{\em Question assumptions} also means {\em question explanations}.
	\begin{itemize}
		\item You: ``I think X is wrong''
		\item Duck: ``How can you be sure?''
	\end{itemize}
	\onslide<2->{
	\linegap
	The {\bf Hypothesis/Test} workflow
	\begin{itemize}
		\item Guess what the root cause is
		\item Write a small test that fails {\em iff} that's right.
		\onslide<3->{
		\begin{itemize}
			\item Not ``change code blindly, see if that fixes it''
			\item Until you know what's wrong, write {\em tests}
		\end{itemize}
		}
	\end{itemize}
	}
	\column{0.25\textwidth}
	\vspace{0.035in}
	\includegraphics[width=\textwidth]{rubberduck.jpg}
	\end{columns}
\end{frame}

@@

	In the ``real world'', and in 400-level classes, failures are not so easy to reproduce.
	\begin{itemize}
		\item Consider: \texttt{input = [rand(), rand(), rand(), rand()]}
		\item Might have failed last week, but never since
	\end{itemize}
	\pause
	\linegap
	Make sure your test cases are {\bf repeatable}
	\begin{itemize}
		\item Getting help (from a TA, from a company) easier when you say ``Here is a failing test'' than ''I once saw a test fail''!
		\item Nonrepeatable tests are often complex
		\begin{itemize}
			\item Use the rest of these techniques to produce a relevant, minimal failing test case.
		\end{itemize}
	\end{itemize}
\end{frame}

@@

\begin{frame}{The Toolbox}
	\begin{columns}[l]
	\column{0.05\textwidth}
	\column{0.5\textwidth}
	\textbf{``Tell Me a Story''}
	\begin{itemize}
		\item Get more detail
		\item Rinse \& repeat
		\pause
		\item {\bf Tell a Story of Time}
		\begin{itemize}
			\item Print debugging
			\item Asserts
		\end{itemize}
		\pause
		\item {\bf Tell a Story of Space}
		\begin{itemize}
			\item Minimize the failing test case
			\item Make small tests that work
		\end{itemize}
		\pause
		\item Draw pictures, collect data, question assumptions, binary search
	\end{itemize}
	\column{0.45\textwidth}
	\pause
	\begin{center}
		\includegraphics[width=0.9\textwidth]{vip1066720.jpg}
	\end{center}
	\end{columns}
\end{frame}

>>>>

% \begin{frame}{Running Example}
%
%	\linegap
%	Sorts OK into the middle and front of list:
%	\begin{itemize}
%		\item \texttt{insert\_sorted} $\quad \hilight{olivegreen}{10} \quad  [20,30] \quad \Rightarrow \quad [\hilight{olivegreen}{10},20,30]$
%		\item \texttt{insert\_sorted} $\quad \hilight{olivegreen}{32} \quad  [16,64] \quad \Rightarrow \quad [16,\hilight{olivegreen}{32},64]$
%	\end{itemize}
%
%	\linegap
%	But not so great at the end:
%	\begin{itemize}
%	\item \texttt{insert\_sorted} $\quad \hilight{red}{100}  \quad [97,98,99] \quad \Rightarrow \quad [97,98,\hilight{red}{100},99]$
%	\end{itemize}
% \end{frame}

@@

	\linegap
		\texttt{~~~~insert\_sorted(list,~element)~\{} \\
			\texttt{~~~~~~~~\hilight{olivegreen}{int}~i~=~\hilight{brickred}{0};} \\
		\texttt{~~~~~~~~\hilight{brown}{while}~(i~<~length(list))~\{} \\
		\texttt{~~~~~~~~~~~~\hilight{brown}{if}~(element~<~list[i])~\{} \\
		\texttt{~~~~~~~~~~~~~~~~\hilight{brown}{break};} \\
		\texttt{~~~~~~~~~~~~\}} \\
		\texttt{~~~~~~~~~~~~i = i + 1;} \\
		\texttt{~~~~~~~~\}} \\
		\texttt{~~~~~~~~insert\_at\_index(list,~element,~i);} \\
		\texttt{~~~~\}} \\
	\pause
	\linegap
	\begin{itemize}
		\item Staring at the code until you see it sometimes works, but is a bad habit.
	\end{itemize}
\end{frame}

@@

\subsection{A Story of Time}

@@

\begin{frame}{Asserts: Stealthier, more concealable prints}
	Adding print statements everywhere is very verbose
	\begin{itemize}
		\item Pro: Get to see everything
		\item Con: Have to look at everything
	\end{itemize}
	\pause
	15-122's C0 ``contracts'' are assertions
	\begin{itemize}
		\item You don't need to be writing C0 to write assert-ful code
		\item Asserts can go anywhere, not just around functions/loops
	\end{itemize}
	\pause
	\linegap

	\texttt{~~~~\hilight{brown}{if}~(something\_that\_should\_never\_happen)~\{} \\
	\texttt{~~~~~~~~print(\hilight{brickred}{"oh~no,~it~happened!"});} \\
	\texttt{~~~~~~~~crash();} \\
	\texttt{~~~~\}} \\
\end{frame}
\begin{frame}{Asserts: Stealthier, more concealable prints}
	Adding print statements everywhere is very verbose
	\begin{itemize}
		\item Pro: Get to see everything
		\item Con: Have to look at everything
	\end{itemize}
	15-122's C0 ``contracts'' are assertions
	\begin{itemize}
		\item You don't need to be writing C0 to write assert-ful code
		\item Asserts can go anywhere, not just around functions/loops
	\end{itemize}
	\linegap

@@

\begin{frame}{Aside: 15-213 ``malloc lab''}
	Many students think: ``I just want to write \texttt{malloc()}. I can do the heap checker later.''
	\pause
	\begin{itemize}
		\item Think of it like a C0 contract that must hold forever.
		\item \textbf{Write the heap checker simultaneously (or even first)!}
			\pause
			\begin{itemize}
				\item Save time/effort debugging everything else
				\item Your debugging skill will level up
				\item Ben jumps up and down for emphasis
			\end{itemize}
	\end{itemize}
\end{frame}

\subsection{A Story of Space}

@@

	%\texttt{~~~~input~=~\hilight{brickred}{[99]};~~~~~\hilight{darkcyan}{//~OK}} \\
	%\texttt{~~~~input~=~\hilight{brickred}{[100,99]};~\hilight{darkcyan}{//~OK}} \\
	%\pause
	%\texttt{~~~~input~=~\hilight{brickred}{[99,100]};~\hilight{darkcyan}{//~Fails}} \\

	\texttt{insert\_sorted} $\quad \hilight{olivegreen}{99}  \quad [~]{\small~~~~} \quad \Rightarrow \quad [\hilight{olivegreen}{99}] \quad$ ~~~~~~(empty list OK) \\
	\texttt{insert\_sorted} $\quad \hilight{olivegreen}{99}  \quad [100] \quad \Rightarrow \quad [\hilight{olivegreen}{99},100] \quad$ (front of list OK) \\
	\pause
	\texttt{insert\_sorted} $\quad \hilight{red}{100}  \quad [99] \quad \Rightarrow \quad [\hilight{red}{100},99] \quad$ (end of list BAD)
\end{frame}

@@

\begin{frame}{\texttt{insert\_sorted}, revisited}
	Now that we know {\em where} and {\em when} the program breaks, let's fix it.
	\begin{itemize}
		\item Raise your hand (don't say anything) when you see what's wrong.
		\item Remember: only fails at the {\em end of the list}.
	\end{itemize}

	\linegap
		\texttt{~~~~insert\_sorted(list,~element)~\{} \\
			\texttt{~~~~~~~~\hilight{olivegreen}{int}~i~=~\hilight{brickred}{0};} \\
		\texttt{~~~~~~~~\hilight{brown}{while}~(i~<~length(list))~\{} \\
		\texttt{~~~~~~~~~~~~\hilight{brown}{if}~(element~<~list[i])~\{} \\
		\texttt{~~~~~~~~~~~~~~~~\hilight{brown}{break};} \\
		\texttt{~~~~~~~~~~~~\}} \\
		\texttt{~~~~~~~~~~~~i = i + 1;} \\
		\texttt{~~~~~~~~\}} \\
		\texttt{~~~~~~~~insert\_at\_index(list,~element,~i);} \\
		\texttt{~~~~\}} \\
\end{frame}
\begin{frame}{\texttt{insert\_sorted}, revisited}
	Now that we know {\em where} and {\em when} the program breaks, let's fix it.
	\begin{itemize}
		\item Raise your hand (don't say anything) when you see what's wrong.
		\item Remember: only fails at the {\em end of the list}.
	\end{itemize}

	\linegap
		\texttt{~~~~insert\_sorted(list,~element)~\{} \\
			\texttt{~~~~~~~~\hilight{olivegreen}{int}~i~=~\hilight{brickred}{0};} \\
			\texttt{~~~~~~~~\hilight{brown}{while}~(\hilight{red}{i~<~length(list)})~\{~\hilight{darkcyan}{//~Should~be~"<="!}} \\
		\texttt{~~~~~~~~~~~~\hilight{brown}{if}~(element~<~list[i])~\{} \\
		\texttt{~~~~~~~~~~~~~~~~\hilight{brown}{break};} \\
		\texttt{~~~~~~~~~~~~\}} \\
		\texttt{~~~~~~~~~~~~i = i + 1;} \\
		\texttt{~~~~~~~~\}} \\
		\texttt{~~~~~~~~insert\_at\_index(list,~element,~i);} \\
		\texttt{~~~~\}} \\
\end{frame}

@@

\begin{frame}{Rubber Duck / Cardboard Dog Debugging}
	\begin{columns}[l]
	\column{0.75\textwidth}
	Hacker folklore says: describe your problem to a duck!
	\begin{itemize}
		\item ``What could I possibly get out of talking to an inanimate object?''
		\pause
		\begin{itemize}
			\item The duck only knows what you tell it
			\item When it doesn't know, it responds with ``WHY?''
		\end{itemize}
	\end{itemize}
	\pause
	\linegap

@@

	Helps figure out what part of the story to start searching at
	}
	\column{0.25\textwidth}
	\includegraphics[width=\textwidth]{rubberduck.jpg}
	\end{columns}
\end{frame}
\begin{frame}{Rubber Duck / Cardboard Dog Debugging}
	\begin{columns}[l]
	\column{0.75\textwidth}
	{\em Question assumptions} also means {\em question explanations}.
	\begin{itemize}
		\item You: ``I think X is wrong''
		\item Duck: ``How can you be sure?''
	\end{itemize}
	\onslide<2->{
	\linegap
	The {\bf Hypothesis/Test} workflow
	\begin{itemize}
		\item Guess what the root cause is
		\item Write a small test that fails {\em iff} that's right.
		\onslide<3->{
		\begin{itemize}
			\item Not ``change code blindly, see if that fixes it''
			\item Until you know what's wrong, write {\em tests}
		\end{itemize}
		}
	\end{itemize}
	}
	\column{0.25\textwidth}
	\vspace{0.095in}
	\includegraphics[width=\textwidth]{rubberduck.jpg}
	\end{columns}
\end{frame}

@@

	In the ``real world'', and in 400-level classes, failures are not so easy to reproduce.
	\begin{itemize}
		\item Consider: \texttt{input = [rand(), rand(), rand(), rand()]}
		\item Might have failed last week, but never since
	\end{itemize}
	\pause
	\linegap
	Make sure your test cases are {\bf repeatable}
	\begin{itemize}
		\item Getting help (from a TA) easier when you say ``Here is a failing test'', not ''I once saw a test fail''!
		\item Nonrepeatable tests are often complex
		\begin{itemize}
			\item Use the rest of these techniques to produce a relevant, minimal, {\em reliable} failing test case.
		\end{itemize}
	\end{itemize}
\end{frame}

@@

\begin{frame}{The Toolbox}
	\begin{columns}[l]
	\column{0.05\textwidth}
	\column{0.5\textwidth}
	\textbf{``Tell Me a Story''}
	\begin{itemize}
		\item Get more detail
		\item Rinse \& repeat
		\pause
		\item {\bf Tell a Story of Time}
		\begin{itemize}
			\item Print debugging
			\item Asserts
		\end{itemize}
		\pause
		\item {\bf Tell a Story of Space}
		\begin{itemize}
			\item Make small tests that work
			\item Minimize the failing test case
		\end{itemize}
		\pause
		\item Draw pictures, collect data, question assumptions, binary search
	\end{itemize}
	\column{0.45\textwidth}
	\pause
	\begin{center}
		\includegraphics[width=0.9\textwidth]{vip1066720.jpg}
	\end{center}
	\end{columns}
\end{frame}
